---
title: "RF Feature Importance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Random Forest Feature Importance

```{r, warning=FALSE, message=FALSE}
library(here)
library(DT)
library(tidyverse)
library(randomForest)
```

```{r read_data}
# external
rf_ext_n2_spl4q_mtry3_ntree100 <- readRDS(here('models/rf/rf_ext_n2_spl4q_mtry3_ntree100.rds'))
rf_ext_n2_spl4q_mtry3_ntree500 <- readRDS(here('models/rf/rf_ext_n2_spl4q_mtry3_ntree500.rds'))
rf_ext_n2_spl4q_mtry3_ntree700 <- readRDS(here('models/rf/rf_ext_n2_spl4q_mtry3_ntree700.rds'))
rf_ext_n2_spl4q_mtry5_ntree100 <- readRDS(here('models/rf/rf_ext_n2_spl4q_mtry5_ntree100.rds'))
rf_ext_n2_spl4q_mtry5_ntree500 <- readRDS(here('models/rf/rf_ext_n2_spl4q_mtry5_ntree500.rds'))
rf_ext_n2_spl4q_mtry5_ntree700 <- readRDS(here('models/rf/rf_ext_n2_spl4q_mtry5_ntree700.rds'))
rf_ext_n2_spl4q_mtry7_ntree100 <- readRDS(here('models/rf/rf_ext_n2_spl4q_mtry7_ntree100.rds'))
rf_ext_n2_spl4q_mtry7_ntree500 <- readRDS(here('models/rf/rf_ext_n2_spl4q_mtry7_ntree500.rds'))
rf_ext_n2_spl4q_mtry7_ntree700 <- readRDS(here('models/rf/rf_ext_n2_spl4q_mtry7_ntree700.rds'))

# internal
rf_int_n2_spl4q_mtry3_ntree100 <- readRDS(here('models/rf/rf_int_n2_spl4q_mtry3_ntree100.rds'))
rf_int_n2_spl4q_mtry3_ntree500 <- readRDS(here('models/rf/rf_int_n2_spl4q_mtry3_ntree500.rds'))
rf_int_n2_spl4q_mtry3_ntree700 <- readRDS(here('models/rf/rf_int_n2_spl4q_mtry3_ntree700.rds'))
rf_int_n2_spl4q_mtry5_ntree100 <- readRDS(here('models/rf/rf_int_n2_spl4q_mtry5_ntree100.rds'))
rf_int_n2_spl4q_mtry5_ntree500 <- readRDS(here('models/rf/rf_int_n2_spl4q_mtry5_ntree500.rds'))
rf_int_n2_spl4q_mtry5_ntree700 <- readRDS(here('models/rf/rf_int_n2_spl4q_mtry5_ntree700.rds'))
rf_int_n2_spl4q_mtry7_ntree100 <- readRDS(here('models/rf/rf_int_n2_spl4q_mtry7_ntree100.rds'))
rf_int_n2_spl4q_mtry7_ntree500 <- readRDS(here('models/rf/rf_int_n2_spl4q_mtry7_ntree500.rds'))
rf_int_n2_spl4q_mtry7_ntree700 <- readRDS(here('models/rf/rf_int_n2_spl4q_mtry7_ntree700.rds'))


```

Feature importance for standardized continuous variables are reported as standardized continuous variables were used as covariates in predictive models. 

```{r}
# convert labels
update_row_names <- function(old_name){
  ifelse(old_name=='Age..scaled.', 'Age', 
  ifelse(old_name=='Female', 'Female',
  ifelse(old_name=='BMI..scaled.', 'BMI', 
  ifelse(old_name=='Disability', 'Disability', 
  ifelse(old_name=='X1st.yr.in.kindergarten', 'Whether first year in kindergarten', 
  ifelse(old_name=='Reading.score..scaled.', 'Read score', 
  ifelse(old_name=='Math.score..scaled.', 'Math score', 
  ifelse(old_name=='Science.score..scaled.', 'Science score', 
  ifelse(old_name=='Lang.at.home', 'Language at home', 
  ifelse(old_name=='Parent.edu', 'Parent education', 
  ifelse(old_name=='HH.income', 'Household income', 
  ifelse(old_name=='SES', 'SES', 
  ifelse(old_name=='Food.security', 'Food security', 
  ifelse(old_name=='HH.members', 'Household member', 
  ifelse(old_name=='Siblings', 'Siblings', 
  ifelse(old_name=='Childcare', 'Childcare', 
  ifelse(old_name=='Home.parent.child.int...scaled.', 'Parent-child interaction', 
  ifelse(old_name=='Reading.together..scaled.', 'Reading together', 
  ifelse(old_name=='School.home.connection..scaled.', 'School-home connection', 
  ifelse(old_name=='Community.violence..scaled.', 'Community violence', 
  ifelse(old_name=='Cultural.heritage..scaled.', 'Cultural heritage', 
  ifelse(old_name=='Approach.learning...parent..scaled.', 'Approach learning (by parent)', 
  ifelse(old_name=='Approach.learning...teacher..scaled.', 'Approach learning (by teacher)', 
  ifelse(old_name=='Attentional.focus..scaled.', 'Attentional focus', 
  ifelse(old_name=='Social.interaction..scaled.', 'Social interaction',
  ifelse(old_name=='Interpersonal..scaled.', 'Interpersonal', 
  ifelse(old_name=='Self.control...parent..scaled.', 'Self-control (by parent)', 
  ifelse(old_name=='Self.control...teacher..scaled.', 'Self-control (by teacher)', 
  ifelse(old_name=='Inhibitory.control..scaled.', 'Inhibitory control', 
  ifelse(old_name=='Community.support..scaled.', 'Community support', 
  ifelse(old_name=='Parent1.Age..scaled.', 'Parent age', 'Child health'
  )))))))))))))))))))))))))))))))
}

update_row_names_list <- function(row_name_list){
  new_names = c()
  for (name in row_name_list){
    new_names = c(new_names, update_row_names(name))
  }
  return(new_names)
}

```

```{r func_imp_table}
importance_table <- function(rf_n2_spl4q){
    rf_n2_spl4q.imp <- setNames(as_tibble(importance(rf_n2_spl4q, scale=FALSE), rownames='x'), c('x', 'low', 'high', 'acc', 'gini')) %>%
        mutate(gini=scales::rescale(gini, to=c(0,100)))
    
    rf_n2_spl4q.imp$x = update_row_names(rf_n2_spl4q.imp$x)
    
    row.names(rf_n2_spl4q$importance) = update_row_names_list(rownames(rf_n2_spl4q$importance))
    
    importance(rf_n2_spl4q, scale=FALSE) %>%
        as.data.frame(rownames=TRUE) %>%
        mutate(across(c(MeanDecreaseAccuracy, MeanDecreaseGini), ~scales::rescale(., to=c(0,100)))) %>%
    DT::datatable(filter='none') %>%
        formatRound(c('Low (2)', 'High (1)', 'MeanDecreaseAccuracy', 'MeanDecreaseGini'), digits=3)
}
```

```{r func_imp_plot_acc}
importance_plot_acc <- function(rf_n2_spl4q){
    rf_n2_spl4q.imp <- setNames(as_tibble(importance(rf_n2_spl4q, scale=FALSE), rownames='x'), c('x', 'low', 'high', 'acc', 'gini')) %>%
        mutate(gini=scales::rescale(gini, to=c(0,100)))
    
    rf_n2_spl4q.imp$x = update_row_names(rf_n2_spl4q.imp$x)
    
    row.names(rf_n2_spl4q$importance) = update_row_names_list(rownames(rf_n2_spl4q$importance))
    
    ps_ext <- rf_n2_spl4q.imp %>%
        pivot_longer(cols=c(acc), names_to='type', values_to='y') %>%
        group_by(type) %>%
        group_map(function(df, key) {
            ggplot(df, aes(reorder(x, y), y)) +
                # geom_point(size=2, pch=21) +
                geom_col() +
                # scale_y_continuous(breaks=function(x) seq(0, ceiling(max(x)/5)*5, 5)) +
                coord_flip() +
                labs(x='', y=paste('Mean decrease in', ifelse(key$type=='acc', 'accuracy', 'gini')))
        })
    ps_ext[[1]]
}
```

```{r func_plot_gini}
importance_plot_gini <- function(rf_n2_spl4q){
    rf_n2_spl4q.imp <- setNames(as_tibble(importance(rf_n2_spl4q, scale=FALSE), rownames='x'), c('x', 'low', 'high', 'acc', 'gini')) %>%
        mutate(gini=scales::rescale(gini, to=c(0,100)))
    
    rf_n2_spl4q.imp$x = update_row_names(rf_n2_spl4q.imp$x)
    
    row.names(rf_n2_spl4q$importance) = update_row_names_list(rownames(rf_n2_spl4q$importance))
    
    ps_ext <- rf_n2_spl4q.imp %>%
        pivot_longer(cols=c(gini), names_to='type', values_to='y') %>%
        group_by(type) %>%
        group_map(function(df, key) {
            ggplot(df, aes(reorder(x, y), y)) +
                # geom_point(size=2, pch=21) +
                geom_col() +
                # scale_y_continuous(breaks=function(x) seq(0, ceiling(max(x)/5)*5, 5)) +
                coord_flip() +
                labs(x='', y=paste('Mean decrease in', ifelse(key$type=='acc', 'accuracy', 'gini')))
        })
    ps_ext[[1]]
}
```


# External
### mtry=3, ntree=100
```{r}
importance_table(rf_ext_n2_spl4q_mtry3_ntree100)
importance_plot_acc(rf_ext_n2_spl4q_mtry3_ntree100)
importance_plot_gini(rf_ext_n2_spl4q_mtry3_ntree100)
```

### mtry=3, ntree=500
```{r}
importance_table(rf_ext_n2_spl4q_mtry3_ntree500)
importance_plot_acc(rf_ext_n2_spl4q_mtry3_ntree500)
importance_plot_gini(rf_ext_n2_spl4q_mtry3_ntree500)
```

### mtry=3, ntree=700
```{r}
importance_table(rf_ext_n2_spl4q_mtry3_ntree700)
importance_plot_acc(rf_ext_n2_spl4q_mtry3_ntree700)
importance_plot_gini(rf_ext_n2_spl4q_mtry3_ntree700)
```

### mtry=5, ntree=100
```{r}
importance_table(rf_ext_n2_spl4q_mtry5_ntree100)
importance_plot_acc(rf_ext_n2_spl4q_mtry5_ntree100)
importance_plot_gini(rf_ext_n2_spl4q_mtry5_ntree100)
```

### mtry=5, ntree=500
```{r}
importance_table(rf_ext_n2_spl4q_mtry5_ntree500)
importance_plot_acc(rf_ext_n2_spl4q_mtry5_ntree500)
importance_plot_gini(rf_ext_n2_spl4q_mtry5_ntree500)
```

### mtry=5, ntree=700
```{r}
importance_table(rf_ext_n2_spl4q_mtry5_ntree700)
importance_plot_acc(rf_ext_n2_spl4q_mtry5_ntree700)
importance_plot_gini(rf_ext_n2_spl4q_mtry5_ntree700)
```

### mtry=7, ntree=100
```{r}
importance_table(rf_ext_n2_spl4q_mtry7_ntree100)
importance_plot_acc(rf_ext_n2_spl4q_mtry7_ntree100)
importance_plot_gini(rf_ext_n2_spl4q_mtry7_ntree100)
```

### mtry=7, ntree=500
```{r}
importance_table(rf_ext_n2_spl4q_mtry7_ntree500)
importance_plot_acc(rf_ext_n2_spl4q_mtry7_ntree500)
importance_plot_gini(rf_ext_n2_spl4q_mtry7_ntree500)
```

### mtry=7, ntree=700
```{r}
importance_table(rf_ext_n2_spl4q_mtry7_ntree700)
importance_plot_acc(rf_ext_n2_spl4q_mtry7_ntree700)
importance_plot_gini(rf_ext_n2_spl4q_mtry7_ntree700)
```

# Internal
### mtry=3, ntree=100
```{r}
importance_table(rf_int_n2_spl4q_mtry3_ntree100)
importance_plot_acc(rf_int_n2_spl4q_mtry3_ntree100)
importance_plot_gini(rf_int_n2_spl4q_mtry3_ntree100)
```

### mtry=3, ntree=500
```{r}
importance_table(rf_int_n2_spl4q_mtry3_ntree500)
importance_plot_acc(rf_int_n2_spl4q_mtry3_ntree500)
importance_plot_gini(rf_int_n2_spl4q_mtry3_ntree500)
```

### mtry=3, ntree=700
```{r}
importance_table(rf_int_n2_spl4q_mtry3_ntree700)
importance_plot_acc(rf_int_n2_spl4q_mtry3_ntree700)
importance_plot_gini(rf_int_n2_spl4q_mtry3_ntree700)
```

### mtry=5, ntree=100
```{r}
importance_table(rf_int_n2_spl4q_mtry5_ntree100)
importance_plot_acc(rf_int_n2_spl4q_mtry5_ntree100)
importance_plot_gini(rf_int_n2_spl4q_mtry5_ntree100)
```

### mtry=5, ntree=500
```{r}
importance_table(rf_int_n2_spl4q_mtry5_ntree500)
importance_plot_acc(rf_int_n2_spl4q_mtry5_ntree500)
importance_plot_gini(rf_int_n2_spl4q_mtry5_ntree500)
```

### mtry=5, ntree=700
```{r}
importance_table(rf_int_n2_spl4q_mtry5_ntree700)
importance_plot_acc(rf_int_n2_spl4q_mtry5_ntree700)
importance_plot_gini(rf_int_n2_spl4q_mtry5_ntree700)
```

### mtry=7, ntree=100
```{r}
importance_table(rf_int_n2_spl4q_mtry7_ntree100)
importance_plot_acc(rf_int_n2_spl4q_mtry7_ntree100)
importance_plot_gini(rf_int_n2_spl4q_mtry7_ntree100)
```

### mtry=7, ntree=500
```{r}
importance_table(rf_int_n2_spl4q_mtry7_ntree500)
importance_plot_acc(rf_int_n2_spl4q_mtry7_ntree500)
importance_plot_gini(rf_int_n2_spl4q_mtry7_ntree500)
```

### mtry=7, ntree=700
```{r}
importance_table(rf_int_n2_spl4q_mtry7_ntree700)
importance_plot_acc(rf_int_n2_spl4q_mtry7_ntree700)
importance_plot_gini(rf_int_n2_spl4q_mtry7_ntree700)
```






