---
title: "AA summary table"
author: 
date: 
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, message=FALSE}
# Load library and dataset needed

library(kableExtra)
library(tidyverse)
library(here)

source(here('code/library/functions.r'))
```

Variables with missingness higher than 30% were excluded for model fitting.
```{r read_data}
dem <- readRDS(here('data/eclsk2011_dem.rds'))
bl <- readRDS('data/bl.rds')
# mdata <- readRDS('data/lcmm_mdata.rds')
mdata_imp <- readRDS('data/lcmm_mdata_imp.rds')

# Filter wide data to students in modeling datasets
# mdata_bl <- bl %>%
#   filter(childid %in% unique(mdata$childid)) # dim: 1250   64
mdata_bl <- bl %>%
  filter(childid %in% unique(mdata_imp$childid)) # dim: 1279   64 imputation: have one or more for spring use imputation

# External group
ext_n2_spl4q <- readRDS(here('data/df_ext_n2_spl4q.rds')) # dim of bl: 1279   68
binom_ext_n2_spl4q <- readRDS(here('models/univariate/binom_ext_n2_spl4q.rds'))

# Internal group
int_n2_spl4q <- readRDS(here('data/df_int_n2_spl4q.rds')) # dim of bl: 1279   68
binom_int_n2_spl4q <- readRDS(here('models/univariate/binom_int_n2_spl4q.rds'))
```


# Comparing baseline characteristics

## Full sample of Asian American children
Summary statistics for the full sample size of interest.
For all continuous variable, the summary statistics were based on original unstandardized values. 

```{r}
xvars <- demvars
xvars[which(xvars=='age_months_sc')] <- 'age_months'
xvars[which(xvars=='p1_pinv11_sc')] <- 'p1_pinv11' # home-child interaction
xvars[which(xvars=='p1_read4_sc')] <- 'p1_read4'  # reading together
xvars[which(xvars=='p2_hscon5_sc')] <- 'p2_hscon5'  # home-school connection
xvars[which(xvars=='com_violence_sc')] <- 'com_violence' # community violence
xvars[which(xvars=='p2_cul2_sc')] <- 'p2_cul2' # cultural heritage
xvars[which(xvars=='x1prnapp_sc')] <- 'x1prnapp' # approach to learn - parent 
xvars[which(xvars=='x1tchapp_sc')] <- 'x1tchapp' # approach to learn - teacher
xvars[which(xvars=='x1attnfs_sc')] <- 'x1attnfs' # attentional focus
xvars[which(xvars=='x1prnsoc_sc')] <- 'x1prnsoc' # social interaction - parent
xvars[which(xvars=='x1tchper_sc')] <- 'x1tchper' # interpersonal - teacher
xvars[which(xvars=='x1prncon_sc')] <- 'x1prncon' # self control - parent 
xvars[which(xvars=='x1tchcon_sc')] <- 'x1tchcon' # self control - teacher
xvars[which(xvars=='x1inbcnt_sc')] <- 'x1inbcnt' # inhibitory control - teacher
xvars[which(xvars=='s4spprt_sc')] <- 's4spprt' # community support 
xvars[which(xvars=='par1age_sc')] <- 'par1age' # parent1 age 
xvars[which(xvars=='p1hscale_sc')] <- 'p1hscale' # health scale
```

```{r}
map(xvars, function(x) {
  if (is.numeric(mdata_bl[[x]])) {
    summarise(mdata_bl, across(all_of(x), ~sprintf('%.2f (%.2f)', mean(., na.rm=T), sd(., na.rm=T)))) %>%
      mutate(x=x) %>%
      rename_with(~paste('val'), .cols=-x)
  } else if (is.factor(mdata_bl[[x]])) {
    count(mdata_bl, !!rlang::sym(x)) %>%
      mutate(across(n, ~sprintf('%i (%.1f%%)', ., . / sum(.) * 100))) %>%
      mutate(across(-n, ~paste(!!x, ., sep=': '))) %>%
      rename_with(~paste('x'), .cols=-n) %>%
      rename(val=n)
  }
}) %>%
  bind_rows() %>%
  select(x, val) %>%
  filter(!str_detect(x, 'FALSE$')) %>%
  kable(booktabs=TRUE, col.names=c('Characteristic', 'Statistic')) %>%
  kable_styling(bootstrap_options=c('striped', 'hover', 'condensed'), full_width=FALSE)

```
note:
`p1_pinv11`: home-child interaction

`p1_read4`: reading together

`p2_hscon5`: home-school connection

`com_violence`: community violence

`p2_cul2`: cultural heritage

`x1prnapp`: approach to learn - parent report 

`x1tchapp`: approach to learn - teacher report

`x1attnfs`: attentional focus

`x1prnsoc`: social interaction - parent report

`x1tchper`: interpersonal - teacher report

`x1prncon`: self control - parent report

`x1tchcon`: self control - teacher report

`x1inbcnt`: inhibitory control - teacher report

`s4spprt`: community support 

`par1age`: parent1 age 

`p1hscale`: health scale



## External group
Descriptive summary stats for high vs. low classes

```{r}
map(xvars, function(x) {
  if (is.numeric(mdata_bl[[x]])) {
    group_by(ext_n2_spl4q$bl, class_lbl) %>%
      summarise(across(all_of(x), ~sprintf('%.2f (%.2f)', mean(., na.rm=T), sd(., na.rm=T)))) %>%
      mutate(x=x) %>%
      rename_with(~paste('val'), .cols=c(-x, -class_lbl))
  } else if (is.factor(mdata_bl[[x]])) {
    group_by(ext_n2_spl4q$bl, class_lbl) %>%
      count(!!rlang::sym(x)) %>%
      mutate(across(n, ~sprintf('%i (%.1f%%)', ., . / sum(.) * 100))) %>%
      mutate(across(-n, ~paste(!!x, ., sep=': '))) %>%
      rename_with(~paste('x'), .cols=c(-n, -class_lbl)) %>%
      rename(val=n)
  }
}) %>%
  bind_rows() %>%
  select(class_lbl, x, val) %>%
  pivot_wider(names_from=class_lbl, values_from=val) %>%
  filter(!str_detect(x, 'FALSE$')) %>%
  kable(booktabs=TRUE) %>%
  add_header_above(c('', 'Class'=2)) %>%
  kable_styling(bootstrap_options=c('striped', 'hover', 'condensed'), full_width=FALSE)

```


Unadjusted logistic regression model. Each characteristic is used as a covariate. The outcome is external class labels from the clustering analysis. 'Low' class is set as the reference group.

All continuous covariates are standardized to $E[X]=0$ and $Var(X)=1$.

```{r}
map(setNames(binom_ext_n2_spl4q, demnames), formatEsts, returnKbl=FALSE) %>%
  bind_rows(.id='covariate') %>%
  {
    idx_hline <- mutate(., i=row_number()) %>%
      group_by(covariate) %>%
      filter(row_number()==n()) %>%
      ungroup() %>%
      pull(i)

    mutate(.,
           or=sprintf('%.2f', or),
           ci=sprintf('[%.2f, %.2f]', cil, ciu),
           pval=case_when(pval < 0.05 ~ sprintf('%.3f*', pval), TRUE ~ sprintf('%.2f', pval))
    ) %>%
      select(covariate, x, or, ci, pval) %>%
      kable(booktabs=TRUE, col.names=c('Covariate', 'x', 'OR', '95% CI', 'p-value')) %>%
      row_spec(idx_hline, hline_after=TRUE) %>%
      column_spec(2, border_right=TRUE) %>%
      collapse_rows(1, valign='middle') %>%
      kable_styling(bootstrap_options=c('hover', 'condensed'), full_width=FALSE)
  }
```


## Internal group
Descriptive summary stats for high vs. low classes

```{r}
map(xvars, function(x) {
  if (is.numeric(mdata_bl[[x]])) {
    group_by(int_n2_spl4q$bl, class_lbl) %>%
      summarise(across(all_of(x), ~sprintf('%.2f (%.2f)', mean(., na.rm=T), sd(., na.rm=T)))) %>%
      mutate(x=x) %>%
      rename_with(~paste('val'), .cols=c(-x, -class_lbl))
  } else if (is.factor(mdata_bl[[x]])) {
    group_by(int_n2_spl4q$bl, class_lbl) %>%
      count(!!rlang::sym(x)) %>%
      mutate(across(n, ~sprintf('%i (%.1f%%)', ., . / sum(.) * 100))) %>%
      mutate(across(-n, ~paste(!!x, ., sep=': '))) %>%
      rename_with(~paste('x'), .cols=c(-n, -class_lbl)) %>%
      rename(val=n)
  }
}) %>%
  bind_rows() %>%
  select(class_lbl, x, val) %>%
  pivot_wider(names_from=class_lbl, values_from=val) %>%
  filter(!str_detect(x, 'FALSE$')) %>%
  kable(booktabs=TRUE) %>%
  add_header_above(c('', 'Class'=2)) %>%
  kable_styling(bootstrap_options=c('striped', 'hover', 'condensed'), full_width=FALSE)

```


Unadjusted logistic regression model. 'Low' class is set as the reference group.

```{r}
map(setNames(binom_int_n2_spl4q, demnames), formatEsts, returnKbl=FALSE) %>%
  bind_rows(.id='covariate') %>%
  {
    idx_hline <- mutate(., i=row_number()) %>%
      group_by(covariate) %>%
      filter(row_number()==n()) %>%
      ungroup() %>%
      pull(i)
    
    mutate(.,
           or=sprintf('%.2f', or),
           ci=sprintf('[%.2f, %.2f]', cil, ciu),
           pval=case_when(pval < 0.05 ~ sprintf('%.3f*', pval), TRUE ~ sprintf('%.2f', pval))
    ) %>%
      select(covariate, x, or, ci, pval) %>%
      kable(booktabs=TRUE, col.names=c('Covariate', 'x', 'OR', '95% CI', 'p-value')) %>%
      row_spec(idx_hline, hline_after=TRUE) %>%
      column_spec(2, border_right=TRUE) %>%
      collapse_rows(1, valign='middle') %>%
      kable_styling(bootstrap_options=c('hover', 'condensed'), full_width=FALSE)
  }
```









